---
title: "Practical Machine Learning Project"
author: "woon kim chun"
date: "September 22, 2015"
output: html_document
---

#Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Data

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 


#Objective

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

```{r loadLibraries}
# load library
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(e1071)
library(randomForest)
set.seed(1)

train.url <-
        "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test.url <- 
        "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# The files are read into memory. Various indicators of missing data (i.e., "NA", "#DIV/0!" and "") are all set to NA so they can be processed.
path <- paste(getwd(),"/", "project", sep="")
train.file <- file.path(path, "machine-train-data.csv")
test.file <- file.path(path, "machine-test-data.csv")


if (!file.exists(train.file)) {
        download.file(train.url, destfile=train.file)
}
if (!file.exists(test.file)) {
        download.file(test.url, destfile=test.file)
}

train.data.raw <- read.csv(train.file, na.strings=c("NA","#DIV/0!",""))
test.data.raw <- read.csv(test.file, na.strings=c("NA","#DIV/0!",""))

```

#Remove NAs column
Remove those columns with NAs

```{r}
# Drop the first 7 columns as they're unnecessary for predicting.
train.data.clean1 <- train.data.raw[,8:length(colnames(train.data.raw))]
test.data.clean1 <- test.data.raw[,8:length(colnames(test.data.raw))]

# Drop colums with NAs
train.data.clean1 <- train.data.clean1[, colSums(is.na(train.data.clean1)) == 0] 
test.data.clean1 <- test.data.clean1[, colSums(is.na(test.data.clean1)) == 0] 

# Check for near zero variance predictors and drop them if necessary
nzv <- nearZeroVar(train.data.clean1,saveMetrics=TRUE)
zero.var.ind <- sum(nzv$nzv)

if ((zero.var.ind>0)) {
        train.data.clean1 <- train.data.clean1[,nzv$nzv==FALSE]
}

```


#Slice the data for cross validation

Divided the training data into two sets. This first is a training set with 70% of the data which is used to train the model. The second is a validation set used to assess model performance.

```{r}
in.training <- createDataPartition(train.data.clean1$classe, p=0.70, list=F)
train.data.final <- train.data.clean1[in.training, ]
validate.data.final <- train.data.clean1[-in.training, ]

```
#Train the Model
The training data-set is used to fit a Random Forest model because it selects important variables automatically . 5-fold cross validation is used when applying the algorithm. A Random Forest algorithm is a way of averaging multiple deep decision trees, trained on different parts of the same data-set, with the goal of reducing the variance. This typically produces better performance at the expense of bias and interpret-ability. The Cross-validation technique assesses how the results of a statistical analysis will generalize to an independent data set. In 5-fold cross-validation, the original sample is randomly partitioned into 5 equal sized sub-samples. a single sample is retained for validation and the other sub-samples are used as training data. The process is repeated 5 times and the results from the folds are averaged.

```{r}
control.parms <- trainControl(method="cv", 5)
rf.model <- train(classe ~ ., data=train.data.final, method="rf",
                 trControl=control.parms, ntree=251)
rf.model

```

#Predict Performance
This model is using the training data which has been used to test against the validation data. Predicted values for the validation data are then compared to the actual values. This allows forecasting the accuracy and overall out-of-sample error, which indicate how well the model will perform with other data.

```{r}
rf.predict <- predict(rf.model, validate.data.final)
confusionMatrix(validate.data.final$classe, rf.predict)

accuracy <- postResample(rf.predict, validate.data.final$classe)
acc.out <- accuracy[1]

overall.ose <- 1 - as.numeric(confusionMatrix(validate.data.final$classe, rf.predict)$overall[1])

```

#Predictive Result
The accuracy of this model is 0.994 and the Overall Out-of-Sample error is 0.0059.

The model is applied to the test data to produce the results.

```{r}
results <- predict(rf.model, test.data.clean1[, -length(names(test.data.clean1))])
results

```

#Decision Tree Visualization view

```{r}
treeModel <- rpart(classe ~ ., data=train.data.final, method="class")
fancyRpartPlot(treeModel)
```

